---
title: "Statistics Denmark StatBank API connection"
output: 
  md_document:
    variant: markdown_github
---

This package connects to the [StatBank](http://www.statistikbanken.dk/statbank5a/) API from [Statistics Denmark](http://www.dst.dk).

This package is in early *BETA* and new changes will most likely not have backward compatibility.

## Installation

You can only install the development version from github, using Hadley Wickham's [devtools](http://cran.r-project.org/web/packages/devtools/index.html) package:

```
if(!require("devtools")) install.packages("devtools")
library("devtools")
install_github("rOpenGov/dkstat")
```

## Examples

Here are a few simple examples that will go through the basics of requesting data from the StatBank
and the structure of the output.  

First, we'll load the library:

```{r, message=FALSE}
library(dkstat)
```

## Meta data

The dst_meta function retrieves meta data from the table you wan't to take a closer look at. It can be used to create the final request, but if you can figure out the structure of the query you can define it yourself.

We'll get some meta data from the [AULAAR table](http://www.statistikbanken.dk/AULAAR). The AULAAR table has net unemployment numbers.

```{r}
aulaar_meta <- dst_meta(table = "AULAAR", lang = "en")
```

The 'dst_meta' function returns a list with 4 objects:
- basics
- variables
- values 
- basic_query

### Basics

Let's see what the basics contains:

```{r}
aulaar_meta$basics
```

There's a table id, a short description, a unit description and when the table was updated.

### Variables

The variables in the list has a short description of each variable as well as the id:

```{r}
aulaar_meta$variables
```

### Values

The values is a list object of all the variable id's you can use to construct your final query:

```{r}
str(aulaar_meta$values)
```

### Basic Query

The basic query is simply a list of the available variables with the first value id for each.

```{r}
aulaar_meta$basic_query
```

## Get data

If you know the table ids from the table you can simply supply the request
through ...

```{r}
aulaar <- dst_get_data(table = "AULAAR", KØN = "TOT", PERPCT = "L10", Tid = 2013,
                       lang = "en", 
                       value_presentation = "ValueAndCode")
str(aulaar)
```


Let's use the basic_query from the dst_meta list to make our first query:

```{r}
aulaar <- dst_get_data(table = "AULAAR", 
                       query = aulaar_meta$basic_query,
                       lang = "en", 
                       value_presentation = "ValueAndCode")
str(aulaar)
```

This is maybe not really what you want, so let's use the basic_query to construct a new query that
might be better. I still want to have the total and percentage unemployed, but I would like all the
observations going back to 1979. I'll now construct the final request, query the StatBank and make a plot.

```{r}
aulaar_meta$basic_query$Tid <- aulaar_meta$values$Tid$id

aulaar <- dst_get_data(table = "AULAAR", 
                       query = aulaar_meta$basic_query, 
                       lang = "en", 
                       format = "CSV",
                       value_presentation = "ValueAndCode")

plot(x = aulaar$TID, 
     y = aulaar$value, 
     main = "Unemployment", 
     xlab = "Year", 
     ylab = "Pct", 
     type = "l")
```

If you want the complete timeseries you can write "*" in the TID variable in the basic_query
or like this:

```{r}
aulaar <- dst_get_data(table = "AULAAR", KØN = "TOT", PERPCT = "L10", Tid = "*",
                       lang = "en", 
                       value_presentation = "ValueAndCode")
str(aulaar)
```

