---
title: "Statistics Denmark StatBank API connection"
output: 
  md_document:
    variant: markdown_github
---

This package connects to the [StatBank](http://www.statistikbanken.dk/statbank5a/) API from [Statistics Denmark](http://www.dst.dk).

This package is in early *BETA* and new changes will most likely not have backward compatibility.

## Installation

You can only install the development version from github, using Hadley Wickham's [devtools](http://cran.r-project.org/web/packages/devtools/index.html) package:

```
if(!require("devtools")) install.packages("devtools")
library("devtools")
install_github("rOpenGov/dkstat")
```

## Examples

Here are a few simple examples that will go through the basics of requesting data from the StatBank
and the structure of the output.  

First, we'll load the library:

```{r, message=FALSE}
library(dkstat)
```

## Meta data

The dst_meta function retrieves meta data from the table you wan't to take a closer look at. It can be used to create the final request, but if you can figure out the structure of the query you can define it yourself.

We'll get some meta data from the [AULAAR table](http://www.statistikbanken.dk/AULAAR). The AULAAR table has net unemployment numbers.

```{r}
aulaar_meta <- dst_meta(table = "AULAAR", lang = "en")
```

The 'dst_meta' function returns a list with 4 objects:
- basics
- variables
- values 
- basic_query

### Basics

Let's see what the basics contains:

```{r}
aulaar_meta$basics
```

There's a table id, a short description, a unit description and when the table was updated.

### Variables

The variables in the list has a short description of each variable as well as the id:

```{r}
aulaar_meta$variables
```

### Values

The values is a list object of all the values in each variable. You use the text column to construct your final query:

```{r}
str(aulaar_meta$values)
```


## Get data

You need to build your query based on the text column that each variable contains in the meta_data$values list.


```{r}
aulaar <- dst_get_data(table = "AULAAR", KØN = "Total", PERPCT = "Per cent of the labour force", Tid = 2013,
                       lang = "en")
str(aulaar)
```


In the above request I don't supply the meta_data to the dst_get_data function, but this is possible
as I will show below. 

Let's query the statbank using more than one value for each variable.

```{r}

folk1_meta <- dst_meta("folk1", lang = "da")

str(dst_get_data(table = "folk1", 
             OMRÅDE = c("Hele landet", "København", "Frederiksberg", "Odense"), 
             STATSB = "Danmark", 
             TID = "*", 
             lang = "da", 
             meta_data = folk1_meta))

```



I can also build a query beforehand and then use the query in the query parameter.

```{r}

my_query <- list(OMRÅDE = c("Hele landet", "København", "Frederiksberg", "Odense"),
                 STATSB = "Danmark",
                 TID = "*")

str(dst_get_data(table = "folk1", query = my_query, lang = "da"))


str(dst_get_data(table = "AUP01", OMRÅDE = c("Hele landet"), TID = "*", lang = "da"))

```

If you want the complete timeseries you can write "*" in the TID variable in your query.

